---
title: "Introduccion a ML. Ejercicios"
author: "Laura Sudupe Medinilla"
date: "17/2/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
# Función para cargar todos los paquetes necesarios
LoadLibraries <- function() {
  myLibraries <- c("faraway")
  invisible(lapply(myLibraries, library, character.only = TRUE))
}
LoadLibraries()
```



EJERCICIOS FARADAY.

1. The dataset teengamb concerns a study of teenage gambling in Britain. Make a
numerical and graphical summary of the data, commenting on any features that
you find interesting. Limit the output you present to a quantity that a busy 
reader would find sufficient to get a basic understanding of the data.

```{r}
# Cargamos el paquete
data <- teengamb
rownames(data) <- c(1:nrow(data))
```

```{r}
#Vamos a ver los datos que lo componen
head(data)
dim(data) #número de filas y columnas
str(data) #estructura de las variables
```
Vemos que tenemos 5 columnas y 47 filas. Todos los valores son numericos, vemos
que la variable sex es categorica

```{r}
data$sex <- factor(data$sex, levels = c(0,1), labels = c("man", "woman"))
summary(data)
```




