---
title: "Estimación del modelo lineal ejercicios"
author: "Laura Sudupe Medinilla"
date: "1/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, echo = FALSE, message = FALSE}
# Función para cargar todos los paquetes necesarios
LoadLibraries <- function() {
  myLibraries <- c("faraway")
  invisible(lapply(myLibraries, library, character.only = TRUE))
}
LoadLibraries()
```

# Ejercicios faraway

## 1.
  The dataset teengamb concerns a study of teenage gambling in Britain. Fit a 
  regression model with the expenditure on gambling as the response and the sex, 
  status, income and verbal score as predictors. Present the output.
```{r}
#ajustamos los datos al modelo
lm <- lm(gamble ~ sex + status + income + verbal, data=teengamb )
lmsum <- summary(lm)
lmsum
```
  (a) What percentage of variation in the response is explained by these 
  predictors?
```{r}
#El coeficiente de determinación R^2 es igual al cuadrado de la correlación del 
#coeficiente. Cuando se expresa en porcentaje, R^2 representa el porcentaje de
#variación en la variable dependiente y puede ser explicado mediante la 
#variación en la variable independiente x mediante la linea de regreión.

(lmsum$r.squared)*100
```
    (b) Which observation has the largest (positive) residual? Give the case 
  number.
```{r}
residuos <- lmsum$residuals
max(residuos)      #valor residuo maximo
which.max(residuos) #observacion con el maximo valor
```
  (c) Compute the mean and median of the residuals.
```{r}
#Han de dar valores entorno al cero, el hecho de que no den es debido a errores
#de calculo
mean(residuos)
median(residuos)
```
  (d) Compute the correlation of the residuals with the fitted values.
```{r}
cor(fitted(lm), resid(lm))
```
  (e) Compute the correlation of the residuals with the income.
```{r}
cor(resid(lm), teengamb$income)
```
  
  (f) For all other predictors held constant, what would be the difference in 
  predicted expenditure on gambling for a male compared to a female?
```{r}
lmsum
lm$coefficients["sex"]
```
  Estos datos nos indican la pendiente de la recta, por lo tanto, la diferencia
  entre "expenditure on gambling in punds per year" es 22.12 "pounds" distinta
  para los hombres que para las mujeres. Menor para las mujeres (sex=1)
  
  
  
  
  
  
  